---
title: "Most deadly terrorist organizations"
author: "Piotr Sobczyk"
date: "March 31, 2016"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
setwd("~/szychta_w_danych/terroryzm/")
library(dplyr)
library(DT)
library(ggplot2)
```

We select only those attacks, that undoubtly were done by terrorist and that were deadly.

```{r}
terrorism <- read.csv("globalterrorism_selected_cols_0615.csv")
terrorism$decade <- floor(terrorism$iyear/10)%%10
terrorism %>% filter(doubtterr==0, nkill>0) -> confirmed.deadly.terrorism
summary(confirmed.deadly.terrorism)
```

What are the most deadly terrorist organizations of all time, which ones organized most attacks?

```{r, echo=FALSE}
confirmed.deadly.terrorism %>% group_by(gname) %>%
	summarise(total_kills=sum(nkill), total_attacks=n()) %>% top_n(20, total_kills) %>%
	arrange(desc(total_kills)) %>% datatable()
```

We can observe that most of the attacks are done by unknown organizations, 
nobody claims responsibility.

For future reference, data is more complete when it comes to the type of weapon used.
This indicates, that database is more or less complete, and on many occasions this is in fact 
difficult to put the blaim on somebody.

```{r, echo=FALSE}
confirmed.deadly.terrorism %>% group_by(weaptype1_txt) %>%
	summarise(total_kills=sum(nkill), total_attacks=n()) %>% top_n(20, total_kills) %>%
	arrange(desc(total_kills)) %>% datatable()
```

It might be interesting to visualize this results by grouping it with respect to 
either a region that given organization is present in, or by its ideology.

#### Unknowns

Let us focus more on those attacks for whom nobody take responsibility - 'unknown's.


where are unknown from? What percentage of attacks is unknown as a percentage of 
all attacks in regions?

```{r, echo=FALSE}
confirmed.deadly.terrorism %>% group_by(region_txt, decade) %>%
	summarise(percent_unknown=mean(gname=='Unknown'),
						count_unknown=sum(gname=='Unknown'),
						kill_unknown=sum(nkill[gname=='Unknown']),
						percent_kill_unknown=kill_unknown/sum(nkill)) %>% ungroup %>% 
	arrange((count_unknown)) %>% datatable()
```

Note that lack of information does not coincide with when attack happend.

```{r, echo=FALSE}
confirmed.deadly.terrorism %>% group_by(region_txt, iyear) %>%
	summarise(percent_unknown=mean(gname=='Unknown')) %>% ungroup -> plot.data
ggplot(plot.data, aes(x=iyear, y=percent_unknown, color=region_txt)) + geom_point() +
	geom_smooth(aes(group=region_txt), se = FALSE)
```

```{r, echo=FALSE}
confirmed.deadly.terrorism %>% group_by(iyear) %>%
	summarise(percent_unknown=mean(gname=='Unknown')) %>% ungroup -> plot.data
ggplot(plot.data, aes(x=iyear, y=percent_unknown)) + geom_point()
```

Some hypothesis:

##### Maybe when source of attack is unknown, then it terrifies people than they start to
act together? Right now, the threat to Western Europe is well known, but it does not bring 
it closer to solution.

##### Nothing is anonymous in well developed parts of the world. 

One reason could be that database is maintained by people from the "west". 

Other is that there is no source of information about attacks
```{r}
confirmed.deadly.terrorism %>% group_by(gname=='Unknown', scite1=="") %>% summarise(n())
```

This is partially true. But maybe attack is more publicized when organization is identified?

Maybe those attacks are really small and that's why nobody investigates them close enough?

```{r}
confirmed.deadly.terrorism %>% group_by(nkill>10, gname=='Unknown') %>% summarise(n())
```

Nope, that's not it.

It seems that the only exlanatory variable here is region. Let us verify it by building 
a logisitc model.

```{r}
fit <- glm((gname=='Unknown')~region_txt+suicide+attacktype1_txt, 
		family = "binomial", data = confirmed.deadly.terrorism)
summary(fit)$coefficients %>% data.frame() %>% datatable()
```

We need to be carefull with interpretation.
